Using best practices for forecasting biodiversity changes in breeding birds

### 2017-08-01 manuscript/draft.Rmd
more edits

---
title: Using best practices for forecasting biodiversity changes in breeding birds
author:
- David J. Harris
- Shawn D. Taylor
- Ethan P. White
bibliography: refs.bib
csl: csl.csl
geometry: margin=1in
fontsize: 12pt
header-includes:
    - \usepackage{times}
    - \usepackage{setspace}
    - \doublespacing
    - \usepackage{lineno}
    - \linenumbers
output:
  pdf_document: default
  html_document: default
---

@@

![](../figures/model_predictions.png)

@@

![](../figures/observer_predictions.png)

@@

![](../figures/scatter.png)

@@

![](../figures/performance_time.png)

@@

![](../figures/observers.png)

>>>>

---
title: Using best practices for forecasting biodiversity changes in breeding birds
author:
- David J. Harris
- Shawn D. Taylor
- Ethan P. White
bibliography: refs.bib
csl: csl.csl
geometry: left=3.5cm, right=3.5cm, top=2.25cm, bottom=2.25cm, headheight=12pt, letterpaper
fontsize: 12pt
header-includes:
    - \usepackage{times}
    - \usepackage{setspace}
    - \doublespacing
    - \usepackage{lineno}
    - \linenumbers
output:
  pdf_document: default
  html_document: default
---

@@

![Example predictions from our six models for a single site. The data points
from 1982 through 2003, connected by solid lines, were used for training all six
models; the remaining points were used for evaluating the models' forecasts. In
each panel, point estimates for each year are shown with lines; the darker
ribbon indicates 1 standard deviation of uncertainty, and the lighter ribbon
indicates 95% confidence intervals. **A.** The time series models models were
trained using one site at a time, based solely on observed richness values
shown. The first two models ("average" and "naive") served as our simple
forecasting baselines. **B.** The environmental models were trained to predict
richness based on elevation, climate, and NDVI; unlike the time series models,
the environmental models' predictions change from year to year as environmental
conditions change.](../figures/model_predictions.png)

@@

![**A.** When all observers are treated the same (black points), predictions can
be poor. **B.** By accounting for systematic differences between observers
(represented by the points' colors), most models can be made more robust to
observer turnover. Note that the "naive" model is already less sensitive to
observer turnover, and does not benefit as much from modeling
it.](../figures/observer_predictions.png)

@@

![Performance of the six models we evaluated when predicting one year into the
future (2004) or ten years into the future (2013).  In general, the time series
models (**A.**) outperformed the environmental models (**B.**), and the accuracy
of the predictions declined as the forecast horizon was
extended.](../figures/scatter.png)

@@

![Performance of all six models over time, according to three metrics. **A.**
Root mean square error (rmse) describes the distance between each model's point
estimates and the observed richness values. Note that the three environmental
models tend to show the largest errors. **B.** Deviance (-2 times the Gaussian
log-likelihood) describes the lack of fit of entire predictive distributions.
Note that the stack of single-species random forest SDMs shows much higher error
than the other models (since its predictions are too confident), and that the
"naive" model's deviance grows relatively quickly (as its predictions become
broader and less informative). **C.** Coverage of a model's 95% confidence
intervals indicates how often the observed values fall inside the predicted
range. For well-calibrated models, the coverage of the 95% confidence intervals
should be 95%. As in the previous panel, the "naive" model's predictive
distribution is too wide (capturing almost all of the data) and the stacked
SDM's predictive distribution is too narrow (missing almost a third of the
observed richness values by 2014).](../figures/performance_time.png)

@@

![Accounting for differences between observers typically has a small impact on
accuracy (especially for the "naive" model), but occasonally changes the
predicted richness values by 10+ species. These changes usually improve the
model's predictions, but not always.](../figures/observers.png)
